---
title: "Exploring the NBA Free Throw Data"
author: "Sam Abbott"
date: "`r Sys.Date()`"
output: 
    html_vignette:
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  
)
```

#Aim

Visualise some aspect of the data you find interesting, e.g., the average number of free throws per period for the regular season and the playoffs.

# Set up
Load analysis packages.

```{r packages}
library(tidyverse)
library(MangoTest)
library(ggridges)
library(plotly)
```

# Data 

Load in the data and explore structure. Data lists the play by play scores in multiple games accross several seasons.

```{r}
str(free_throws)
```

Split out teams, and scores in each game.

```{r split scores and teams}
tidy_free_throws <- free_throws %>% 
  separate(game, c("team_1", "team_2"), " - ") %>% 
  separate(score, c("score_1", "score_2"), " - ") %>% 
  separate(end_result, c("end_result_1", "end_result_2"), " - ")
```



Seperate game wide information, play information and scores information. Convert to factors as appropriate.

```{r game-plays-info}
games_df <- tidy_free_throws %>% 
  select(game_id, team_1, team_2, playoffs,
         season, end_result_1, end_result_2) %>% 
  unique %>% 
  mutate(draw = ifelse(end_result_1 == end_result_2, "Yes", "No")) %>% 
  gather(key = "order", value = "team", team_1, team_2) %>% 
  separate(order, c("tmp", "order"), "_") %>% 
  select(-tmp) %>% 
  gather(key = "order2", value = "end_result", 
         end_result_1, end_result_2) %>% 
  separate(order2, c("tmp", "order2"), "result_") %>% 
  select(-tmp) %>% 
  filter(order == order2) %>% 
  select(-order2) %>% 
  mutate_at(.vars = vars(order, team, season, playoffs),
            .funs = funs(factor(.))) %>% 
  mutate(end_result = as.numeric(end_result)) %>% 
  mutate(season = factor(season, levels = rev(paste0(2006:2015, " - ", 2007:2016))))

plays_df <- tidy_free_throws %>% 
  select(game_id, period, play, player, shot_made, time) %>% 
  mutate(player = factor(player))

scores_df <- tidy_free_throws %>% 
  select(game_id, score_1, score_2, period, time) %>% 
    gather(key = "order", value = "score", score_1, score_2) %>% 
  separate(order, c("tmp", "order"), "_") %>% 
  select(-tmp) %>% 
  mutate(order = factor(order)) %>% 
  mutate(score = as.numeric(score))
```

Add game outcome variable.

```{r game-outcome}
games_df <- games_df %>% 
  group_by(game_id) %>%
  arrange(desc(end_result), .by_group = TRUE) %>% 
  mutate(outcome = c("Won", "Lost")) %>% 
  ungroup() %>% 
  mutate(outcome = ifelse(draw %in% "Yes", "Draw", outcome)) %>% 
  mutate(outcome = factor(outcome)) %>% 
  select(-draw)
```

Summarise each dataset in turn. No draws in the dataset - online search suggests NBA games are played until a team wins.

```{r sum-df}
summary(games_df)
summary(plays_df)
summary(scores_df)
```

# Visualisations

## Comparision of Final Scores over Multiple Seasons


```{r mult-season-final-scores}
games_df %>%
  ggplot(aes(x = end_result, y = season, fill = outcome)) +
  geom_density_ridges(alpha = 0.6) +
  theme_minimal() +
  scale_fill_viridis_d() +
  guides(fill = guide_legend(title = "Game Outcome")) +
  labs(x = "Final Score",
       y = "Season",
       title = "Distribution of NBA Final Scores",
       subtitle = "By Game Outcome from 2006 until 2016",
       caption = "By Sam Abbott, for Mango Solutions. Source: NBA Free Throws")
```

## Times at Which Points were Scored During Play

```{r}
scores_df %>% 
  left_join(games_df %>% 
              select(game_id, order, team, outcome, season),
            by = c("game_id", "order")) %>%  
  mutate(period = factor(period)) %>% 
  group_by(team, period, game_id) %>% 
  arrange(desc(score), .by_group = TRUE) %>% 
  slice(1) %>%
  ungroup %>% 
  ggplot(aes(x = score, y = period,
             fill = outcome)) +
  geom_density_ridges(alpha = 0.6) +
  scale_fill_viridis_d() +
  theme_minimal() +
  guides(fill = guide_legend(title = "Game Outcome")) +
  labs(x = "Score",
       y = "Period",
       title = "Distribution of NBA Scores during Play",
       subtitle = "By Game Outcome for all Teams: 2006 - 2016",
       caption = "By Sam Abbott, for Mango Solutions. Source: NBA Free Throws")
```

## Conversion Rates of Free Throws by Player


```{r}
conv_player_df <- plays_df %>% 
  left_join(games_df %>% 
              select(game_id, season) %>% 
              unique,
            by = "game_id") %>% 
  group_by(player, season) %>% 
  summarise(conversion = mean(shot_made), shots = n()) %>% 
  mutate(conversion = round(conversion, digits = 3)) %>% 
  ungroup

conv_player_df %>% 
  plot_ly(y = ~conversion*100, x = ~shots, 
          text = ~player,
          frame = ~season) %>%
  add_markers() %>% 
  layout(yaxis = list(title = "Conversion Rate", range = c(0, 110))) %>% 
  layout(yaxis = list(ticksuffix = "%")) %>% 
  layout(xaxis = list(title = "Free Throws")) %>% 
  layout(title = "NBA Free Throw Conversion Rate") %>% 
  animation_opts(2000, redraw = FALSE, easing = "elastic") %>% 
  animation_slider(currentvalue = list(prefix = "Season: ")) %>% 
  hide_legend()
```
